\documentclass[a4paper,DIV12]{scrartcl}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{url}
\usepackage{color}
\usepackage{listings}
\usepackage[pdftex]{hyperref}
\graphicspath{{pics/}}
\include{commands}
\bibliographystyle{apalike}
\parindent 0pt
\author{
Thomas G{\"u}nther\thanks{Leibniz Institute for Applied Geophysics, Hannover (Germany)} 
\ \& \ 
Carsten R{\"u}cker\thanks{Institute of Geology and Geophysics, University of Leipzig (Germany)}
}
\title{GIMLi --\\ Geophysical Inversion and Modelling Library \\-- programmers tutorial}
\newcommand{\cw}[1]{{\bf\texttt{#1}}}
\newcommand{\invert}{invert}
\newcommand{\command}[1]{\par{\tt \$ #1}}
\newcommand{\file}[1]{{\tt #1}}
\newcommand{\sperre}{\,\rule{1ex}{2ex}}
\newcommand{\mycomment}[2]{\sperre{\em #1: #2}\sperre}

\lstloadlanguages{[Visual]C++,[ISO]C++,Python}
\lstset{language=[ISO]C++}
\lstdefinelanguage{C++}{morekeywords={RVector,RMatrix,ModellingBase}}
%\lstset{numbers=left,numberstyle=\tiny}
%\lstset{basicstyle=\small,keywordstyle=\color{red}}
\lstset{basicstyle=\sffamily\small,keywordstyle=\bfseries}
\lstset{frame=shadowbox,commentstyle=\color{blue}}
\lstset{morekeywords={RVector,RMatrix,ModellingBase,RSparseMapMatrix,SparseMatrix}}

\begin{document}
\maketitle
\begin{abstract}
GIMLi is a C++ class library for solving inverse and forward problems in geophysics.
It was build in order to make inversion available to various forward modelling routines.
Template programming algorithms are used to create a purely mathematical framework for solving physics problems.
The inversion can apply various minimisation and regularization schemes, different transformation functions a region approach for sophisticated regularization and the chance to incorporate parametric and structural inversion.

In this tutorial we like to show the programmer how to work with GIMLi by means of different examples.
A simple curve-fitting is used to show how different forward operators can be incorporated.
Some small 1d examples demonstrate the use of different parameterisation and the partitioning of data and model.
The region technique can be used for different target parameters or geological units.

Joining different inversion runs is one of the key issues of ongoing research.
We show three different joint inversion (JI) types: one-parameter JI, petrophysical JI and structural JI.
Finally the time-lapse inversion framework is presented and applied to ERT data using different algorithms.
\end{abstract}

\clearpage
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{firstexample}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{concepts}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{enhanced}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{jointinversion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Inversion templates}
There are certain tasks that go beyond a classical inversion scheme but still are method-independent.
They can be formulated more generally in order to be applied to a wider range of applications.
There are several cases
\begin{description}
\item[Roll-along inversion:] Geophysical data are often acquired along profiles. Since numerical work usually goes by $N^2$, it can be much more efficient to do the work piece-wise. However, the continuity must be ensured.
\item[Joint inversion:] As described above, the different joint inversions do not have to be programmed individually. Furthermore, more than two methods can be coupled.
\item[Time-lapse inversion:] Although there are several approaches for efficient inversion along the time axis, they are method-independent and can be formulated generally.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Roll-along inversion}
\sperre{BLA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Joint inversion}
Assume we have inversions invA, invB and invC, which (or parts of which) are to be coupled.

\subsubsection*{Block inversion}
Block inversion: mesh1d containing thickness and several parameters.
\sperre{TO BE IMPLEMENTED}

\subsubsection*{Structural coupling}
As described in section \ref{sec:structjoint}, the structural coupling consists mainly some preparation steps and a main iteration for the coupling that does individual runs and coupling.

Generally, there can be two possibilities for each inversion: (i) to couple the whole inversion model (as above) or (ii) to couple only one specified region.
For example, structures in an aquifer are to be imaged by cross-hole ERT and GPR. Whereas for the latter it is sufficient to restrict to the aquifer, for ERT the unsaturated zone and a clay layer must be taken into account.

For more than 2 inversions the coupling must be generalized: One can imaging a chain (or ring) or a star scheme coupling. 
Whereas in the first the inversions are coupled pair-wise, in the latter the  constraint weights are mixed. 

\begin{lstlisting}
SCCInversion SCC;
SCC.append( invA );
SCC.append( invB, regionnumber );
SCC.run(); 
\end{lstlisting}

Options: type of coupling, e.g. IRLS scheme or self-defined function. Ring or star connection. Terminating criterion.

\sperre{TO BE IMPLEMENTED}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time lapse inversion}\label{sec:timelapse}
\subsubsection*{Strategies}\label{sec:tlstrategies}
\subsubsection*{Example}
%DC1d infiltration??

\bibliography{gimli} 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{appendix}
\include{appendix}
\end{appendix}

\end{document}
